Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-23 WSSSPE2_report/WSSSPE2_review.tex
Section 3.1: Changes due to feedback from Thaney.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes} \label{sec:keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{\footnotesize{Lead: Choi. Live notes: \href{http://tinyurl.com/q6ew45v}{Google Notes} and \href{http://tinyurl.com/pd9tmh3}{here}}.
%https://docs.google.com/document/d/1-BxkYWDQ6nNNBXBStUL0xcKF9qCTlEALwf928J_MemI/edit?usp=sharing
Abstracts: \href{http://tinyurl.com/mnenzms}{WSSSPE2}. Slides: \href{http://tinyurl.com/las8tsq}{here}.}

@@

Thaney started noting the unfortunate fact that many current systems suffer the
unintended consequence of creating friction that hinders users, despite
designers' original purposes to do good. An example is the National Cancer
Institute's caBIG. A total of \$350 million was spent, including more than \$60
million for management. More than $70$ tools were created, but caBIG is still
seen as a failure\footnote{\url{http://tinyurl.com/maf6dz2}}. Those that had the
least investment were the most used; the most invested software were the least
utilized.
\begin{comment}
(Kaitlin didn't really say why? Were the tools used? Did they immediately
bit-rot? Was their development unsustained?)-should ask question about this. Is
there a caBIG report available? Quotes from Andrea Califano/Joe Gray, looks like
there is: \url{tinyurl.com/qdodbo7}.
%http://www.informationweek.com/architecture/report-blasts-problem-plagued-cancer-research-grid/d/d-id/1097068

@@

Thaney emphasized that for efficient reproducible open research, we would need
research tools (e.g., software repositories), social capital (e.g., incentives),
and capacity (e.g., training and mentorship). Our systems would need to
communicate with each other. A point was made by a member of the auidence that
as systems become less monolithic, it often becomes harder to sustain the links between
them\footnote{See, for example, \url{http://tinyurl.com/l76tba2}.}.
%http://www.slideshare.net/jameshowison/scientific-software-sustainability-and-ecosystem-complexity
%(but does Anon Grizzly have other cites/info for that?).
In contrast, works licensed by Creative Commons typically do not 
have dependencies (a book, a photo, an artwork).

Thaney introduced Mozilla's Software Discovery Index
(SDI)\footnote{\url{http://softwarediscoveryindex.org/report/}}, which
facilitates more efficient scientific research. SDI identifies scientific
software by archiving and standardizing metadata for software and hence help
connect both developers and users. Ted Habermann in the audience commented that
if the metadata is minimal, it would be less onerous for data providers, but
more burdensome for users---it could be challenging to keep a balance between
what have to be captured and what would be ideal if we do not want to lose user
engagement as in the case of the old Harvard Dataverse, finding often only the
first four fields of three pages of metadata were filled out.

The speaker concluded her talk urging the audience to design scientific software
with the general community, not an individual, in mind; and to design to unlock
latent potential of our systems. In addition, she encouraged everyone to rethink
how we reward researchers and support roles. Lastly, she cautioned the community
to be mindful of jargon or semantics traps.

@@

We do not recognise these hundreds of thousands of researchers because we think
they are the long tail and forget that they are the mainstream and we are the
elite minority.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes\todo{\footnotesize{Lead: Choi. \href{http://tinyurl.com/q6ew45v}{Google Notes} and \href{http://tinyurl.com/pd9tmh3}{here}}.
%https://docs.google.com/document/d/1-BxkYWDQ6nNNBXBStUL0xcKF9qCTlEALwf928J_MemI/edit?usp=sharing
\href{http://tinyurl.com/mnenzms}{Abstracts}.  \href{http://tinyurl.com/las8tsq}{Slides}.}
} \label{sec:keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Thaney started noting the unfortunate fact that many current systems suffer the
unintended consequence of creating friction that hinders users, despite
designers' original purposes to do good. An example is the National Cancer
Institute's caBIG. A total of \$350 million was spent, including more than \$60
million for management. More than $70$ tools were created, but caBIG is still
seen as a failure\footnote{\url{http://tinyurl.com/maf6dz2}}. Those that had the
least investment were the most used; the most invested software were the least
utilized.

\begin{comment}
(Kaitlin didn't really say why? Were the tools used? Did they immediately
bit-rot? Was their development unsustained?)-should ask question about this. Is
there a caBIG report available? Quotes from Andrea Califano/Joe Gray, looks like
there is: \url{tinyurl.com/qdodbo7}.
%http://www.informationweek.com/architecture/report-blasts-problem-plagued-cancer-research-grid/d/d-id/1097068

@@

Thaney emphasized that for efficient reproducible open research, we would need
research tools (e.g., software repositories), social capital (e.g., incentives),
and capacity (e.g., training and mentorship). Our systems would need to
communicate with each other. A point was made by a member of the audience that
as systems become less monolithic, it often becomes harder to sustain the links
between them\footnote{See, for example, \url{http://tinyurl.com/l76tba2}.}.
%http://www.slideshare.net/jameshowison/scientific-software-sustainability-and-ecosystem-complexity
%(but does Anon Grizzly have other cites/info for that?).
%In contrast, works licensed by Creative Commons typically do not 
%have dependencies (a book, a photo, an artwork).

Thaney spoke about Mozilla's work around code citation, through a collaboration
and prototype crafted between Mozilla, Github, figshare and Zenodo. This work
was presented at a closed meeting in May 2014 at the National Institutes of
Health (NIH) around these issues, sparking a conversation from that meeting
around what a \emph{Software Discovery
Index}\footnote{\url{http://softwarediscoveryindex.org/report/}} might look
like. The meeting included a number of publishers, researchers, and those behind
major scientific software efforts such as Bioconductor, Galaxy, and
nanoHUB. 
%facilitates more efficient scientific research. SDI identifies scientific
%software by archiving and standardizing metadata for software and hence help
%connect both developers and users.
Ted Habermann in the audience commented that if the metadata is minimal, it
would be less onerous for data providers, but more burdensome for users---it
could be challenging to keep a balance between what have to be captured and what
would be ideal if we do not want to lose user engagement as in the case of the
old Harvard Dataverse, finding often only the first four fields of three pages
of metadata were filled out.

The speaker concluded her talk urging the audience to design scientific software
with the general community, not an individual, in mind; and to design to unlock
latent potential of our systems. In addition, she encouraged everyone to rethink
how we reward researchers and support roles.
%Lastly, she cautioned the community to be mindful of jargon or semantics traps.

@@

We do not recognize these hundreds of thousands of researchers because we think
they are the long tail and forget that they are the mainstream and we are the
elite minority.
